Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-04 EconGrowth.Rnw
Now creating a table of parameters from Cobb-Douglas with energy fits.

print("US"); print(cobbDouglasPredictions("US", "Q"))
print("UK"); print(cobbDouglasPredictions("UK", "Q"))
print("JP"); print(cobbDouglasPredictions("JP", "Q"))
print("CN"); print(cobbDouglasPredictions("CN", "Q"))
print("ZA"); print(cobbDouglasPredictions("ZA", "Q"))
print("SA"); print(cobbDouglasPredictions("SA", "Q"))
print("IR"); print(cobbDouglasPredictions("IR", "Q"))
print("TZ"); print(cobbDouglasPredictions("TZ", "Q"))
print("ZM"); print(cobbDouglasPredictions("ZM", "Q"))

@@

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){
  modelCD <- cobbDouglasModel(countryName)
  # Checks validity of the model. AIC stands for Akaike's Information Criterion.
  aicCD  <- AIC(modelCD, k=2)
  summaryCD <- summary(modelCD) # Gives the nls summary table.
  ciCD <- confint(modelCD, level = ciLevel); ciCD # Calculates confidence intervals for the CD model.
  # Calculate beta and its confidence interval and report it.
  alpha <- as.numeric(coef(modelCD)["alpha"])
  beta <- 1.0 - alpha
  beta.est <- deltaMethod(modelCD, "1 - alpha"); beta.est # Estimates beta and its standard error (SE).
  # Now calculate a confidence interval on beta
  dofCD <- summaryCD$df[2]; dofCD # Gives the degrees of freedom for the model.
  tvalCD <- qt(ciHalfLevel, df = dofCD); tvalCD
  betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCD <- data.frame(lambda = coef(modelCD)["lambda"], alpha = coef(modelCD)["alpha"], beta = beta, gamma = 0)
  row.names(estCD) <- "CD"
  # The [1] subscripts pick off the lower confidence interval
  lowerCD <- data.frame(lambda = ciCD["lambda","2.5%"], alpha = ciCD["alpha", "2.5%"], beta = betaCICD[1], gamma = NA) 
  row.names(lowerCD) <- "-95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCD <- data.frame(lambda = ciCD["lambda","97.5%"], alpha = ciCD["alpha", "97.5%"], beta = betaCICD[2], gamma = NA)
  row.names(upperCD) <- "+95% CI"
  # Now create the data for a table.
  dataCD <- rbind(upperCD, estCD, lowerCD)
  return(dataCD)
}

############
# Creates a row for the Cobb Douglas parameters table
##
cobbDouglasCountryRow <- function(countryName){
  dataCD <- cobbDouglasData(countryName)
  # Create CI ranges
  out <- cbind(dataCD["-95% CI", "lambda"], dataCD["CD", "lambda"], dataCD["+95% CI", "lambda"],
               dataCD["-95% CI", "alpha"],  dataCD["CD", "alpha"],  dataCD["+95% CI", "alpha"],
               dataCD["-95% CI", "beta"],   dataCD["CD", "beta"],   dataCD["+95% CI", "beta"])
  return(out)
}

@@

############################
# Aggregates the Cobb-Douglas results into a big data table
##
cobbDouglasParamsTableNoEnergy <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas (without energy) for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}
@

<<CD Params Table, results='asis'>>=
print(cobbDouglasParamsTableNoEnergy(), caption.placement="top", sanitize.colnames.function = identity)
@

<<CD_Params_Graph, fig.width=figWidth, fig.height=3.5, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=FALSE>>=
createCDParamsGraph <- function(){
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))

@@

>>>>

@@

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string, one of "Q", "X", or "U"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName, energyType){
  modelCD <- cobbDouglasModel(countryName, energyType)
  summaryCD <- summary(modelCD) # Gives the nls summary table.
  ciCD <- confint(modelCD, level = ciLevel); ciCD # Calculates confidence intervals for the CD model.
  dofCD <- summaryCD$df[2] # Gives the degrees of freedom for the model.
  tvalCD <- qt(ciHalfLevel, df = dofCD)
  if (missing(energyType)){
    #Cobb-Douglas without energy
    alpha <- as.numeric(coef(modelCD)["alpha"])
    beta <- 1.0 - alpha
    beta.est <- deltaMethod(modelCD, "1 - alpha") # Estimates beta and its standard error (SE).
    # Calculate beta and its confidence interval and report it.
    # Now calculate a confidence interval on beta
    betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
    # Combine all estimates and their confidence intervals into data frames with intelligent row names
    estCD <- data.frame(lambda=coef(modelCD)["lambda"], alpha=coef(modelCD)["alpha"], beta=beta, gamma=0)
    row.names(estCD) <- "CD"
    # The [1] subscripts pick off the lower confidence interval
    lowerCD <- data.frame(lambda=ciCD["lambda","2.5%"], alpha=ciCD["alpha", "2.5%"], beta=betaCICD[1], gamma=NA) 
    row.names(lowerCD) <- "-95% CI"
    # The [2] subscripts pick off the lower confidence interval
    upperCD <- data.frame(lambda=ciCD["lambda","97.5%"], alpha=ciCD["alpha", "97.5%"], beta=betaCICD[2], gamma=NA)
    row.names(upperCD) <- "+95% CI"
  } else {
    #Cobb-Douglas with energy
    a <- as.numeric(coef(modelCD)["a"])
    b <- as.numeric(coef(modelCD)["b"])
    print(a)
    print(b)
    lambda <- as.numeric(coef(modelCD)["lambda"])
    alpha <- a
    beta <- b - a
    gamma <- 1.0 - b
    # Report results with SE
    beta.est <- deltaMethod(modelCD, "b-a") # Reports results for beta, because beta = b - a.
    gamma.est <- deltaMethod(modelCD, "1-b") # Reports results for gamma, because gamma = 1 - b.
    # Now calculate confidence intervals.
    betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE) # Gives the confidence interval on beta.
    gammaCICD <- with(gamma.est, Estimate + c(-1.0, 1.0) * tvalCD * SE) # Gives the confidence interval on gamma.
    # Combine all estimates and their confidence intervals into data frames with intelligent row names
    estCD <- data.frame(lambda = lambda, alpha = alpha, beta = beta, gamma = gamma)
    row.names(estCD) <- "CDe"
    # The [1] subscripts pick off the lower confidence interval
    lowerCD <- data.frame(lambda=ciCD["lambda","2.5%"], alpha=ciCD["a", "2.5%"], beta=betaCICD[1], gamma=gammaCICD[1])
    row.names(lowerCD) <- "- 95% CI"
    # The [2] subscripts pick off the lower confidence interval
    upperCD <- data.frame(lambda=ciCD["lambda","97.5%"], alpha=ciCD["a", "97.5%"], beta=betaCICD[2], gamma=gammaCICD[2])
    row.names(upperCD) <- "+ 95% CI"
  }
  # Now create the data for a table.
  dataCD <- rbind(upperCD, estCD, lowerCD)
  return(dataCD)
}

############
# Creates a row for the Cobb Douglas parameters table
##
cobbDouglasCountryRow <- function(countryName, energyType){
  dataCD <- cobbDouglasData(countryName, energyType)
  if (missing(energyType)){
    out <- cbind(dataCD["-95% CI", "lambda"], dataCD["CD", "lambda"], dataCD["+95% CI", "lambda"],
                 dataCD["-95% CI", "alpha"],  dataCD["CD", "alpha"],  dataCD["+95% CI", "alpha"],
                 dataCD["-95% CI", "beta"],   dataCD["CD", "beta"],   dataCD["+95% CI", "beta"])
  } else {
    out <- cbind(dataCD["-95% CI", "lambda"], dataCD["CDe", "lambda"], dataCD["+95% CI", "lambda"],
                 dataCD["-95% CI", "alpha"],  dataCD["CDe", "alpha"],  dataCD["+95% CI", "alpha"],
                 dataCD["-95% CI", "beta"],   dataCD["CDe", "beta"],   dataCD["+95% CI", "beta"],
                 dataCD["-95% CI", "gamma"],  dataCD["CDe", "gamma"],  dataCD["+95% CI", "gamma"])
  }
  return(out)
}

@@

############################
# Aggregates the Cobb-Douglas results into a big data table. No energy.
##
cobbDouglasParamsTableNoEnergy <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas (without energy) for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

############################
# Aggregates the Cobb-Douglas results into a big data table. With energy.
##
cobbDouglasParamsTableWithEnergy <- function(energyType){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow, energyType=energyType))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ",
                        " ", "$\\gamma$",  " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption=paste("Cobb-Douglas (with ", energyType, ") for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}
@

<<CD_Params_Table_No_E, results='asis'>>=
print(cobbDouglasParamsTableNoEnergy(), caption.placement="top", sanitize.colnames.function = identity)
@

<<CD_Params_Graph, fig.width=figWidth, fig.height=3.5, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph <- function(){
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))

@@

<<CD_Params_Table_Q, results='asis'>>=
print(cobbDouglasParamsTableWithEnergy("Q"), caption.placement="top", sanitize.colnames.function = identity)
@
